---
title: "American’s And Financial Satisfaction"
author: "Jason Eyerly"
date: "September 20, 2014"
output: html_document
---

  This document is a project proposal for my Coursera class: Data Analysis & Statistical Inference. As a person who is actively involved in the Financial Markets, I hypothesized that there may be a correlation between the satisfaction of Americans and the percentage return of the S&P500 the year prior. 

  My reasoning for this is that the $SPX can often be used as an economic indicator, and may in turn reflect how Americans are doing financially.

```{r, echo=FALSE}
require(Quandl)
require(lubridate)
require(zoo)
require(xts)

myGSS <- load(url("http://bit.ly/dasi_gss_data"))

year <- gss$year
finSat <- gss$satfin

relativeTable <- data.frame(year, finSat)
relativeTable <- subset(relativeTable, year > "1988" & !is.na(finSat))


spReturns <- Quandl("SANDP/ANNRETS", trim_start="1970-01-11", 
                    trim_end="2012-12-31", authcode="nwy3a_Gmd7TSS9fVirxT", 
                    collapse="annual")

percentChange <- spReturns$"Total Return Change"

spReturns$"Year Ending" <- format((spReturns$"Year Ending"), "%Y")
spReturns$"Year Ending" <- as.numeric(spReturns$"Year Ending")
spReturns$"Year Ending" <- spReturns[,1] + 1 #the following year

combined <- merge(relativeTable, spReturns, by.x = "year", by.y = "Year Ending")
names(combined)[6] <- "percentChange"

finalResults <- data.frame(combined$year, combined$finSat, combined$percentChange)
names(finalResults)[1] <- "Year"
names(finalResults)[2] <- "FinancialSatisfaction"
names(finalResults)[3] <- "PercentChange"

finalResults$PercentChange <- finalResults$PercentChange * 100
```

Boxplot Summarizing Results:

```{r, echo=FALSE}
boxplot(finalResults$PercentChange ~ finalResults$FinancialSatisfaction, xlab = "Financial Satisfaction", ylab = "Percent Change")
summary(finalResults$PercentChange)
table(finalResults$FinancialSatisfaction)
```

As you can see, the data is left skewed, and is unimodal. This left skew implies that the mean was dragged down by a greater number of years with a low or even negative return.
